Validating FF-Replan vs.\@ a State-Of-The-Art Probabilistic Planner in Translated Real World Domains

### 2014-04-08 FinalPaper/paper.tex
ff replan done, fixed a label
New domain description

\usepackage{xcolor}
\newcommand\myworries[1]{\textcolor{red}{#1}}
\newcommand\todo[1]{\textcolor{blue}{#1}}

@@

% first the title is needed
\title{Realistically Interesting Domains for Probabilistic Planners}

% a short form should be given in case it is too long for the running head
\titlerunning{Real World Domains}

@@

@@

\section{Background}
\ref{sec:background}

@@

FF-Replan \cite{FFReplan}, a replanner based on determinization techniques and a classical planner, surprisingly outperformed other participating MDP solvers in the IPPC-04 and by doing so changed the course of the planning competition entirely. By relying on search and removing all probabilistic information from the problem, FF-replan functions with enormous speed and efficiency. This method has proven to outperform MDP solvers, the state-of-the-art probabilistic planners for uncertainty of that time. Since then, Little et al. has argued for \emph{probabilistically interesting} domains. By introducing finite horizon MDPs in IPPC-11, state of the art planners have been adapted to cope with large branching factors to approximate infinite planning problems. Because these "planning problems" can often be solved using a smaller horizon $h < H$ as introduced in Section \ref{sec:background}, they represent optimization problems rather than planning problems, favoring current state of the art planners over replanners. This section introduces the shift of the planners towards optimization  by discussing FF-replan and a current state of the art probabilistic planner \textsc{GOURMAND} and its predecessors.

\myworries{misschien moet dat laatste zinnetje ergens anders?}

\subsection{FF-replan}
\label{sec:ffreplan}
 FF-Replan generates a deterministic planning problem by determinizing the input domain at the start of each planning iteration. This classical deterministic planning problem allows for the use of the classical FF planner \cite{Hoffmann01theff}. The FF planner then returns an ordered state-action plan for the deterministic problem of the current state. This plan is executed until an unexpected state is encountered in the following manner:

\begin{enumerate}
	\item Start planning at $s_0$
	\item Determinizes the input domain
	\item Use FF on the classical planning problem
	\item Executes ordered state-action plan until
	\item Unexpected state is encountered or
	\begin{itemize}
		\item Start planning at $s_t$ at time t
		\item Repeat step 2 - 6
	\end{itemize}
	\item Goal state or dead-end is reached
\end{enumerate}

FF-Replan has some obvious shortcomings. It removes all probabilistic information from the problem when determinizing the domain. The success of the planner is therefore fully reliable on the chosen determinization heuristic. As FF-replan uses no roll-out policy or future simulation, it has no way of avoiding \emph{avoidable} dead-ends. This explains why FF-replan performs poorly in domains containing dead-ends.

For a planner to be succesfull in finite-horizon MDPs it needs to be scalable and sensitive to an expected reward maximization, the evaluation criteria of these domains. This caused for  LRTDP planners to emerge.

%----------------------------------------------------------------------------------------

\subsection{The rise of LRTDP planners}

\myworries{Still need to introduce LRTDPs better and the reason why they emerged, ook UCT introduceren?}

\emph{Labeled Real-Time Dynamic Programmers} (LRTDPs) operate a series of trials starting from $s_0$.

\todo{Make this list more technical}
\begin{itemize}
	\item Trials start at $s_0$
	\item Choose best action greedily in the current state based on current $V$ value function
	\item Perform Bellman backup on the current state
	\item Sample the outcome as the chosen action
	\item Transition to the new state
	\item Repeat until a goal, dead end or converged state is reached
	\item After trial, special convergence check on all states in the trial is performed to prove the convergence of these values
\end{itemize}

Adapting LRTDP to a finite horizon MDP \emph{M(H)} or \emph{LRTDP-FH} requires for each trial to start at ($s_0$) and run for at most $H$ time steps. The resulting states would translate to the goal states in a goal-oriented MDP. \\

%However, it's anytime performance to meet the demands of a dynamic environment is not very good, creating the need for a more sophisticated approach \cite{kolobov2012glutton}.

A \emph{Labeled Reverse Real-Time Dynamic Programmer} (LR\textsuperscript{2}TDP) uses \emph{reverse iterative deepening} and LRTDP-FH to incrementally solve a sequence of MDPs $M(1)$, $M(2)$, \ldots , $M(H)$ in order by relying on the fact that a successful policy can be produced by solving an MDP-FH $M(H)$ for a smaller horizon $h < H$. A LR\textsuperscript{2}TDP\@ iterates on the distance from the horizon and allows for reusing precomputed state-action plans from converged states. However when dealing with large branching factors Bellman computations tend to become prohibitively expensive. 
\todo{nog een lijstje schrijven van hoe het werkt en waarom er verbetering nodig is}

%\begin{itemize}
%	\item In many MDPs, the optimal policy $M(h)$ for some $h < H$ is near optimal for M(H), it thus may have a solution when stopped prematurely
%	\item it's trials are very short because backchaining from the goal leads to a lots of convergence
%	\item It optimally memorizes a large fraction of augmented states to save memory
%	\item a large fraction of state space may be unreachable from $s_0$ and it avoid storing information it does not need
%	\item it should however use a heuristic to treat states that has never been encountered when stopped prematurely, could be a fall back or a default action.
%\end{itemize}

\subsection{GLUTTON}

\textsc{GLUTTON} introduces an off-line algorithm based on LR\textsuperscript{2}TDP. The \emph{Max-Reward} heuristic used to evaluate \todo{the what?} relies on knowing the maximum reward $R_{max}$ any action can yield in a state or an upper bound on it. In addition to LR\textsuperscript{2}TDP, it introduces the following mechanisms to optimize for large branching factors:

\todo{rewrite this list}
\myworries{is this really necessary?}
\begin{itemize}
	\item Sub sampling the transition function, is not enough to decrease the amount of computations yet
	\item Separating out Natural Dynamics: to reuse variable samples when generator successors for multiple actions, by separating the actions' effects into natural dynamics (\emph{exogenous}), those due to the action itself (\emph{pure effects}) and those do a mix of both.
	\item Caching the transition function samples: as LR\textsuperscript{2}TDP populates the state-value hash table lazily, Glutton uses  this memory as a cache for samples from the transition function, so it doesn't have to regenerate the action when encountering the same $(s,h)$ pair. When full, it chooses state-action pars for eviction ans replacement randomly.
	\item Default policies: Glutton uses 3 tricks to determine a policy when stopped prematurely. It compares the optimal action from a $h'$ with a simple cyclic policy or a complete random policy and chooses the best of these based on the reward.
\end{itemize}

%---------------------------------------------------------------------
%During IPPC-11 1 out of 5 planners used a LRTDP-based system and managed to outperform three other planners using the dominant algorithm for probabilistic planning: UCT. This caused to wonder whether LRTDP planners could outperform UCT based systems.  UCT forces an expert to guess a non-shortsighted lookahead value for which it should be able to converge on encountered states. Choosing this value can have very bad consequences and its existence causes the UCT to require some input, taking away from its autonomicity. At the IPPC-11 GLUTTON was an offline LRTDP based planner that came second. As PROST employed UCT in an online manner, GOURMAND is an online adaptation of the GLUTTON algorithm to test whether LRTDP performs better if it is employed online.
%
%UCT picks an action based on it current quality estimate and an \emph{exploration term} which forces the planner to choose actions that have rarely been tried in the past. This makes UCT suitable for a wider scale of scenarios but also implies an unreliable termination condition.

Because real world domains contain Natural Dynamics that introduce large branching factors, offline planning is often a waste of time and memory space since there is a vast amount of states the planner may never reach in a lifetime. Such problems are far better solved using online planners by finding an optimal policy $\tau$* or an approximation of this polict for the \emph{current} state, executing it and repeating. This is done by choosing a certain lookahead value $L$ and solving the MDP \emph{M\textsubscript{s}(L)} to solve \emph{M\textsubscript{0}(H)} where L $<$ H and then select \emph{$\pi$*\textsubscript{L}(s, L)} using the evaluation criteria. This lookahead value is determined using the per-process time constraint.\todo{elaborate on the per-process time constraint}

\subsection{GOURMAND}
\textsc{GOURMAND}, an online adaptation of the \textsc{GLUTTON} planner answers these remarks by adding a lookahead value $L$. \todo{still finish this}

% always makes an informed choice in any state where it ends up when the time T allocated for solving the MDP runs out because of the online fashion it uses LRTDP. GOURMAND first divides the time among all remaining epochs, determines the largest lookahead L for which it can almost certainly solve any remaining epoch if T is assigned.  It then checks if it can solve even for a larger lookahead L by taking away a small configuration time from future epochs  while still guaranteeing that it can solve for Lookahead L in each of them. This was GOURMAND adaptively picks a good lookahead value at each decision epoch given the per-process constraint T. This can only be achieved by knowing how long it takes for LRTDP to solve a state for certain lookahead values. It uses LR\textsuperscript{2}TDP to collect this data. This version of LRTDP is more efficient because of the early convergences and it allows measuring how long on average the solving of a task takes. GOURMAND uses the same engineering optimizations to to handle evaluate actions, iterating over the successors of s under a by the Bellman Backups. GOURMAND eventually outperforms both GLUTTON and PROST, the winner of the IPPC-11\cite{keller2012prost}.

>>>>

@@

% first the title is needed
\title{Validating FF-Replan vs.\@ a State-Of-The-Art Probabilistic Planner in Translated Real World Domains}

% a short form should be given in case it is too long for the running head
\titlerunning{FF-Replan vs Probabilistic Planner in Real World Domains}

@@

\section{Domain Description}

Unmanned Aerial Vehicles (UAVs), or commonly called \emph{Drones} are aircrafts
without a pilot and often autonomously controlled. Drones are widely used for
military as well as for civil uses, including inspecting pipelines, counting
wildlife, agriculture, filming or delivering supplies to inaccessible areas.
As many uses cases there are, there are many forms of UAVs, bigger fixed wing
aircrafts, or small helicopter or quad copter like aircrafts that can hover at
a certain position in the air.

We are particularly interested in monitoring, for example crops or livestock,
and surveillance UAVs that work together to monitor a large area. Each UAV
monitors a part of the total area, and if it has monitored something, it has
to communicate this to a base station. This can't be done directly due to the
large distance, instead it has to go through neighboring UAVs.

Communicating the message from the sensing UAV to the base station can be
expressed as a probabilistic planning problem. Initially the sensing UAV
has the package that has to be sent, and the goal is to bring this package to
the base station. Each time the package is passed to the neighboring UAV,
there is a possibility the package is not received correctly. The bigger the
distance, the higher the probability the delivery failed.

%% TODO maybe UAVs can be on or of too, so it's forced to use a further UAV.
%% TODO what's also interesting is that Amazon.com uses drones for package deliveries

@@

\section{Background}

@@

\subsection{FF-Replan}

At the first international probabilistic planning competition IPPC-04, most entries were Markov Decision Processes. The winner, FF-Replan, however was based on deterministic planning techniques.

FF-Replan is an action selection algorithm for online planning in probabilistic domains. Given a domain description, a goal and an initial state as a probabilistic planning problem, FF-Replan generates a deterministic planning problem. It determinizes the input domain at the start of the planning into a deterministic different domain. They consider two methods of determinization:
\begin{enumerate}
\item \emph{Single-outcome determinization}: selects one outcome for each probabilistic construct using a variety of heuristics. The performance strongly depends on the methods for choosing the outcomes.
\item \emph{All-outcomes determinization}: considers every probabilistic outcome as a distinct deterministic action. For each possible outcome FF-Replan generates one seperate action for each effect. For nested probabilities, the all-outcomes procedure is applied recursively. This implies that FF-Replan will cease selecting actions whenever it enters a dead state. However when given no information, all probabilistic effects will be considered as equal which is a potential weakness.
\end{enumerate}

After determinization FF-Replan then uses the deterministic FF planner to compute an ordered plan for the generated deterministic problem. This process is repeated when countering an unexpected state until a goal state is reached. It maintains a partial state-action mapping using a hash-table which is initially empty. Whenever FF-Replan encounters a state which is not in the table, it determinizes the problem and synthesizes a plan using FF. It then simulates the plan according to the deterministic action definitions resulting in a state-action sequence whose pairs are put in the hash table. The first action of the plan is then executed in the environment, returning a new state. This partial policy is this produces in an online fashion. It would be valuable to look into incremental planning approaches that don't create a deterministic plan from scratch whenever an unexpected state is encountered or reusing previous planning effort.

FF-Replan has some obvious shortcomings as it first determinizes the input domain, removing all probabilistic information from the problem and then synthesizes a plan. If an unexpected state should occur whilst executing, the planner replans in the same determinization. It also does not consider multiple potential effects if an action.

By relying on search, FF-Replan could function with enormous speed and efficiency. However when dealing with difficult probabilistic problems attempting to span the spectrum of complexity ranging from deterministic problems to highly stochastic problems, FF-Replan is expected to come short. Because FF-Replan is developed hoping to inspire extensions and insight into the approach and planning domains, the authors are hoping to achieve this.

\subsection{GOURMAND}

The IPPC-11 introduces more realistic MDPs to accommodate real-world systems with large branching factors cause by \emph{natural dynamics} and a limited time for the planner to come up with a policy. For this reason planners needed to be scalable and sensitive to the expected reward maximization. Glutton introduces an off-line algorithm based on LR\textsuperscript{2}TDP and relies on the fact that a successful policy can be produced by solving an MDP $M(H)$ for a smaller horizon $h$. Incrementally solving the sequence $M(1)$, $M(2)$ forms the basis of LR\textsuperscript{2}TDP\@. By iterating on the distance from the horizon, data can be reused (and is stored in the cache until the hash-table is full). The max-reward heuristic finds the largest horizon value $h'$ for which already a heuristic value $V$ is estimated. The  computation time remains much lower because it runs quickly into states that converged while solving $M(h-1)$. The max-reward computation relies on knowing the maximum $R_{max}$ any action can yield in any state or an upper bound on it.

\subsubsection{LRTDP}
Using goal oriented MDPs, the \emph{Labelled Real-Time Dynamic Programmer} (LRTDP) operates a series of trials starting from $s_0$.

\begin{itemize}
	\item Trials start at $s_0$
	\item Choose best action greedily in the current state based on current $V$
	\item Perform  Bellman backup on the current state
	\item Sample the outcome as the chosen action
	\item Transition to the new state
	\item Repeat until a goal, dead end or converged state is reached
	\item After trial, special convergence check on all states in the trial is performed to prove the convergence of these values
\end{itemize}

\subsubsection{LRTDP-FH}

Adapting LRTDP to a finite horizon MDP \emph{M(H)} requires to let each trial start at ($s_0$) and run for at most H time steps. A steps $H$ steps away from $s_0$ would translate to goals in a goal-oriented MDP. However, it's anytime performance to meet the demands of a dynamic environment is not very good, creating the need for a more sophisticated approach \cite{kolobov2012glutton}.

\subsubsection{LR\textsuperscript{2}TDP}

LR\textsuperscript{2}TDP uses \emph{reverse iterative deepening} and LRTDP-FH to solve a sequence of MDPs $M(1)$, $M(2)$, \ldots , $M(H)$, in that order. It first decides how to act optimally in ($s_0$, 1), equivalent to solving M(1) , then ($s_0$, 2), equivalent to solving $M(2)$, eventually solving $M(H)$ or running out of time halting a $M(h')$ where $h' < H$. This incremental effort allows for reusing state values computed in previous states, so it memorizes all values and solves for smaller horizons by backchaining from the goal LRTDP-FH iterates over search depth while LR\textsuperscript{2}TDP iterates over the distance from the horizon and is likely to be better because:

\begin{itemize}
	\item In many MDPs, the optimal policy $M(h)$ for some $h < H$ is near optimal for M(H), it thus may have a solution when stopped prematurely
	\item it's trials are very short because backchaining from the goal leads to a lots of convergence
	\item It optimally memorizes a large fraction of augmented states to save memory
	\item a large fraction of state space may be unreachable from $s_0$ and it avoid storing information it does not need
	\item it should however use a heuristic to treat states that has never been encountered when stopped prematurely, could be a fall back or a default action.
\end{itemize}

\subsubsection{GLUTTON}

Glutton uses the \emph{Max-Reward} heuristic that relies on knowing the maximum reward $R_{max}$ any action can yield in a state or an upper bound on it. Large branching factors induce that the computation of the Bellman Backups becomes prohibitively expensive. Glutton uses the following mechanisms to optimize for large branching factors:

\begin{itemize}
	\item Subsampling the transition function, is not enough to decrease the amount of computations yet
	\item Separating out Natural Dynamics: to reuse variable samples when generator successors for multiple actions, by separating the actions' effects into natural dynamics (\emph{exogenous}), those due to the action itself (\emph{pure effects}) and those do a mix of both.
	\item Caching the transition function samples: as LR\textsuperscript{2}TDP populates the state-value hash table lazily, Glutton uses  this memory as a cache for samples from the transition function, so it doesn't have to regenerate the action when encountering the same $(s,h)$ pair. When full, it chooses state-action pars for eviction ans replacement randomly.
	\item Default policies: Glutton uses 3 tricks to determine a policy when stopped prematurely. It compares the optimal action from a $h'$ with a simple cyclic policy or a complete random policy and chooses the best of these based on the reward.
\end{itemize}
%---------------------------------------------------------------------
During IPPC-11 1 out of 5 planners used a LRTDP-based system and managed to outperform three other planners using the dominant algorithm for probabilistic planning: UCT. This caused to wonder whether LRTDP planners could outperform UCT based systems.  UCT forces an expert to guess a non-shortsighted lookahead value for which it should be able to converge on encountered states. Choosing this value can have very bad consequences and its existence causes the UCT to require some input, taking away from its autonomicity. At the IPPC-11 GLUTTON was an offline LRTDP based planner that came second. As PROST employed UCT in an online manner, GOURMAND is an online adaptation of the GLUTTON algorithm to test whether LRTDP performs better if it is employed online.

UCT picks an action based on it current quality estimate and an \emph{exploration term} which forces the planner to choose actions that have rarely been tried in the past. This makes UCT suitable for a wider scale of scenarios but also implies an unreliable termination condition.

Because real world domains contain Natural Dynamics that introduce large branching factors, offline planning is often a waste of time and memory space since there is a vast amount of state the planner may never reach in a lifetime. Such problems are far better solved using online planners by finding a $\tau$* or an approximation for the current state, executing it and repeating. This is done by choosing a certain lookahead  L and solving MDP \emph{M\textsubscript{s}(L)} to solve \emph{M\textsubscript{0}(H)} where L $<$ H and then select \emph{$\pi$*\textsubscript{L}(s, L)} recommended by the optimal policy. This lookahead value is determined using the per-process time constraint.

\subsubsection{GOURMAND}
In contrast to GLUTTON, GOURMAND always makes an informed choice in any state where it ends up when the time T allocated for solving the MDP runs out because of the online fashion it uses LRTDP. GOURMAND first divides the time among all remaining epochs, determines the largest lookahead L for which it can almost certainly solve any remaining epoch if T is assigned.  It then checks if it can solve even for a larger lookahead L by taking away a small configuration time from future epochs  while still guaranteeing that it can solve for Lookahead L in each of them. This was GOURMAND adaptively picks a good lookahead value at each decision epoch given the per-process constraint T. This can only be achieved by knowing how long it takes for LRTDP to solve a state for certain lookahead values. It uses LR\textsuperscript{2}TDP to collect this data. This version of LRTDP is more efficient because of the early convergences and it allows measuring how long on average the solving of a task takes. GOURMAND uses the same engineering optimizations to to handle evaluate actions, iterating over the successors of s under a by the Bellman Backups. GOURMAND eventually outperforms both GLUTTON and PROST, the winner of the IPPC-11\cite{keller2012prost}.
